### 2021-11-15 inst/main/linear_test.Rmd
All files
files

---
title: "linear_test"
author: "Anubhav Gupta"
date: "16/12/2020"
output: html_document
---

```{r message=FALSE}

library(dplyr)
library(ggplot2)

set.seed(123)
m <- 4
c <- 5
npoints <- 3001
x <- seq(0, 300, length.out = npoints)
epsilon <- rnorm(npoints, mean = 0, sd = 50)
y <- m * x + c + epsilon

dd_raw <- data.frame(x = x, y = y)
ggplot(dd_raw) +
  geom_point(aes(x = x, y = y)) +
  geom_vline(xintercept = c(100, 200), color = "red")

dd_lower <- dd_raw %>%
  filter(x <= 100)

dd_mid <- dd_raw %>%
  filter(x >= 100 & x <= 200)

dd_upper <- dd_raw %>%
  filter(x >= 200 & x <= 300)

lm_all <- lm(y ~ x, data = dd_raw)
lm_lower <- lm(y ~ x, data = dd_lower)
lm_mid <- lm(y ~ x, data = dd_mid)
lm_upper <- lm(y ~ x, data = dd_upper)

summary(lm_all)
summary(lm_lower)
summary(lm_mid)
summary(lm_upper)
```

We observe that the standard error in the estimation of intercept increases. The standard error in the slope remains approximately equal for the three ranges (lower, mid and upper) of x values.