### 2022-02-28 manuscript/ms_C2/ms_C2.Rmd
Update ms_C2.Rmd

---
title: How much gut content data is required to predict trophic interactions?
authors:
- affiliation: University of Zurich
  department: Department of Evolutionary Biology and Environmental Studies
  email: anubhav.gupta@ieu.uzh.ch
  location: 8057 Zurich, Switzerland
  name: Anubhav Gupta
  thanks: Corresponding author
- affiliation: University of Essex
  department: School of Life Sciences
  email: e.ogorman@essex.ac.uk
  location: CO4 3SQ Colchester, UK
  name: Eoin O' Gorman
- affiliation: University of Zurich
  department: Department of Evolutionary Biology and Environmental Studies
  email: owen.petchey@ieu.uzh.ch
  location: 8057 Zurich, Switzerland
  name: Owen L. Petchey
geometry: margin=1in
abstract: |

  1)  Presence absence data of species interactions such as gut content data is used to parameterise food web models such as the allometric diet  breadth model (ADBM), thereby predicting interactions. Collecting and analysing these data from the field is, however, an expensive and time consuming task. Therefore, it is crucial to know how much gut content data is required to parameterise food web models with high accuracy and high precision.
  2)  We apply to this problem a recently developed approach, namely Bayesian computation to parameterise the ADBM, and true skill statistics to measure the goodness of fit, and do so while varying the amount of gut content data is used in the parameterisation. 
  3)  We estimated the minimum number of gut content data required for seven different food webs that resulted in 95% of the maximum true skill statistics achieved by the ADBM when all available gut content data was used. We found that incomplete gut content data can be used to parameterise the ADBM with the lowest amount of gut content data being 28% of the available gut content data.
  4)  These results suggest that one need not collect such a large quantity of gut content data to predict a food web structure using a food web model, thereby reducing sampling effort considerably, while having little effect on precision or accuracy of predictions.

keywords:
  - gut content data
  - ADBM
  - food web accuracy
  - food web prediction
header-includes:
   - \usepackage{lineno}
   - \linenumbers
   - \usepackage {amsmath}
   - \setlength\parindent{24pt}
   - \usepackage{setspace}\doublespacing
# indent: true
mainfont: Times New Roman
bibliography: bibliography.bib
biblio-style: biblatex
output:
  # word_document:
  #   fig_caption: yes
  #   number_sections: yes
  rticles::arxiv_article:
    keep_tex: true
---

\*\*AG: My comment starts with "**AG:"

# Alternate titles

-   Predicting trophic interactions using incomplete gut content data

-   Effect of the amount of gut content data on the accuracy and precision of food web prediction

# Introduction

Knowledge about the trophic interactions in a food web is crucial in ecology for purposes ranging from identifying keystone species [@jordanKeystoneSpeciesFood2009] to quantifying robustness of a food web to species extinctions [@dunneNetworkStructureBiodiversity2002]. This has led to the development of numerous food web models [@allesinaGeneralModelFood2008; @cohenStochasticTheoryCommunity1985; @gravelInferringFoodWeb2013; @petcheySizeForagingFood2008; @tamaddoni-nezhadConstructionValidationFood2013]. Along with inferring missing links in an observed food web, the food web models are also used for ecological forecasting [@hattabForecastingFinescaleChanges2016; @lindegrenEcologicalForecastingClimate2010] and for understanding the underlying mechanism governing the interactions in a food web [@ogormanSimpleModelPredicts2019].

Although food web models are constructed using prior theory about the factors that determine trophic interactions, empirical data about interactions is required to parameterise a model. For example, @petcheySizeForagingFood2008 used presence-absence information about trophic interactions to parameterise the allometric diet breadth model and thereby predict species interactions. These empirical information about interactions can come from diverse set of methods such as gut content analysis [@peralta-maraverStructureDynamicsStability2017], stable isotope ratio analysis of tissues [@laymanCanStableIsotope2007], experimentation [@warrenSpatialTemporalVariation1989], DNA metabarcoding of gut contents or faeces [@roslinUseDNABarcodes2016] and literature research [@grayJoiningDotsAutomated2015; @cohenSoilInvertebratesChemistry2014; @goldwasserConstructionAnalysisLarge1993a].

<!-- **OP: The following three sentences may not be necessary here, or could perhaps go elsewhere.** All of these methods can provide information useful to infer the diet of a consumer, however different methods can result in different levels of uncertainty in the inferred diet (citreqd). A single gut content of a predator describes what that individual has consumed recently which can be used to infer the diet of the predator [@eitzingerTestingValidityFunctional2018; @omalleyEffectsFoodWeb2017; @dixonAssessingDietNorth2017; @mcclain-countsTrophicStructureMesopelagic2017b; @peralta-maraverStructureDynamicsStability2017]. Stable isotope data when combined with mixing models can be used to determine which prey items are most likely fed upon by a predator [@kadoyaIsoWebBayesianIsotope2012; @crawfordApplicationsStableIsotope2008]. -->

Each of these sources of information about trophic interactions has shortcomings. Stable isotope ratio analysis of the organism's tissue does gives indirect information of the diet of that organism via revealing approximate trophic position of that species in the food web. Although mixing models can be used to determine what prey items are most likely fed upon by a predator, this results in uncertainty in the estimates [@kadoyaIsoWebBayesianIsotope2012; @crawfordApplicationsStableIsotope2008]. Experimentation (e.g. feeding trials) may create unrealistic conditions where a predator is confined in a small space with a potential "prey" and is effectively "forced" into eating it (citreqd). DNA metabarcoding could have many issues such as failure to relate OTUs to species, or inability to distinguish secondary predation, i.e. the DNA signature of the prey consumed by the prey consumed by the primary item found in the predator gut. Another approach to infer missing links is literature research which involves assuming that a link will occur because another author described it in some other system, or for related predator/prey species, e.g. from the same genus or family [@grayJoiningDotsAutomated2015; @cohen2014; @goldwasser1993]. This typically results in a food web that has false positives--links included when in reality no link would occur (citreqd). @nielsen2018 have shown that gut content analysis method has a better match with real diet when compared to other methods. \*\*AG: Working on this paragraph.

Acquiring food web data from gut content analysis is time consuming and expensive (reference). In part this is due to the perception that many gut contents must be collected and analysed in order to be confident that the majority of possible trophic links among species have been observed. Furthermore, considerable time, resource investment, and expertise is required to process the collected samples. Hence, there is of considerable importance to know the minimum number of gut contents and of which consumer species that are required to parameterise a food web model with high accuracy and high precision, so that one knows when is a good time to stop collecting any further empirical data on the food web.

Therefore, the key question we are interested in answering is how much presence-absence information, in the form of gut content data, is required to infer food web structure from a food web model with high accuracy and high precision. In other words, how many samples of gut content should one collect from the field to parameterise a food web model? To answer this question, we use the presence-absence information from gut content data to parameterise the allometric diet breadth model (ADBM) thereby predicting trophic interactions in seven different food webs using rejection approximate Bayesian computation, where the fit of the model was measured using true skill statistics. We computed the true skill statistics of the predicted food webs for different number of predator guts to calculate the minimum number of gut content data to infer food web structure. Our study provides a guideline on how many gut content data is required to predict food web structure using a food web model.

```{r echo = FALSE, warning = FALSE}
# options(knitr.kable.NA = '')
# library(readxl)
# library(knitr)
# library(kableExtra)
# fw_table <- read_excel("../table/data_summary_fil.xlsx")
# 
# kable(fw_table, format = "latex",
#       caption = "\\label{fig:tab_1}Studies showing the type of data used to infer trophic interactions in a food web",
#       escape = FALSE) %>%
#     kable_styling(latex_options="scale_down")

```

# Materials and Methods

In the upcoming sections, we present the empirical food webs, the allometric diet breadth model (ADBM), and the gut content data used to infer the trophic interactions. We also give a detailed account of using partial gut content data to parameterise the ADBM using rejection approximate Bayesian computation (ABC). We assessed model predictions using the true skill statistic and various structural food web properties for comparison across food webs.

## The Empirical Food Webs

In our study, we used food webs for which gut content data is available at an individual level and data that is or that we could make FAIR [OP: cite this https://www.go-fair.org/fair-principles/]. Presence and absence interactions of food web data can be aggregated in different ways [@gilljamSeeingDouble2011]. A common way of aggregating food web data is the taxonomic approach, in which nodes in the food web are taxa (e.g. species), or life stages of taxa, and one species is said to feed on another species if at least one individual of the latter is found in at least one gut of the former. Another approach is based on size, rather than taxonomy, and in this the nodes in the food web are size classes. Here, a feeding link occurs between two size classes if at least one prey item within a size class was found in the gut of another size class, irrespective of their taxonomy.

In our study, we aggregate the gut content data on the basis of size class because the food webs constructed using this approach were better predicted by the ADBM compared to the taxonomic approach [@woodward2010], and this is because the ADBM is based on foraging rules defined using the body sizes of organisms [@petcheySizeForagingFood2008].

Our study is based on the assumption that the food web constructed using the complete gut content data is the true food web. However, this might not be the case because the links are still under-sampled for many nodes [@woodwardIndividualBasedFoodWebs2010]. This could result in a lower true skill statistic (TSS) of the predicted food web from the ADBM, which can result in a biased estimate of the number of gut content data [OP: this is unclear].

```{r echo=FALSE, warning=FALSE}
library(readxl)
library(knitr)
library(kableExtra)
fw_table <- read_excel("../table/fw_table.xlsx")
kable(fw_table, format = "latex",
      caption = "\\label{fig:tab_1}Information about the empirical food webs. **AG: Working on this.", escape = FALSE, align = "l") %>%
  kable_styling(font_size = 7, latex_options="scale_down", row_label_position = 1) %>%
  landscape(margin = "1cm") %>%
  column_spec(1, width = "3cm") %>%
  column_spec(2, width = "8em") %>%
  column_spec(7, width = "8em") %>%
  column_spec(8, width = "8em") %>%
  column_spec(9, width = "8em")

```

<!-- ### Broadstone Stream -->

[OP: I like these sections and suggest to include them for other food webs also.]

<!-- Broadstone Stream (51$^\circ$ 05' N 0$^\circ$ 03' E; 120 m above sea-level) is a second-order tributary of the River Medway in south-east England. It is dominated by invertebrates as the acidity of the stream (pH 4.7-6.6) excludes fish. The stream consists of 25 common invertebrate species [@woodwardQuantificationResolutionComplex2005]. Among the predators, the three large species are *Cordulegaster boltonii* Donovan, *Sialis fuliginosa* Pict. and *Plectrocnemia conspersa* [Curtis] and the three small species are the larvae of the tanypod midges *Macropelopia nebulosa* [Meigen], *Trissopelopia longimana* [Staeger], and *Zavrelimyia barbatipes* [Kieffer]. Broadstone Stream food web is one of the most completely described food webs available for any system [@schmid-arayaImportanceMeiofaunaFood2002; @hildrewChapterSustainedResearch2009; @layerFoodWebStructure2010; and @woodwardQuantificationResolutionComplex2005a]. -->

<!-- ### Celtic Sea -->

<!-- The Celtic Sea is an area of continental shelf bordered by Ireland, the UK and the Bay of Biscay. Precise sampling locations and dates were not given in the Barnes et al. (2008) dataset, from where the data used in this chapter were extracted, so we pooled data over the whole time period and locations to capture general patterns. Only locations consistently sampled through the 1987--2001 time series were used (Blanchard et al., 2005). The feeding links of fishes in the Celtic Sea have been described in a published global dataset of individual predator and prey body sizes and taxonomy (Barnes et al., 2008): in total, 1988 feeding events from 29 predator species were included in the food web presented here. The original stomach contents data were collected from dissections carried out on board research vessels during the annual surveys carried out by the Centre for Environment, Fisheries and Aquaculture Science (Cefas) (Pinnegar et al., 2003). Predator and prey length were recorded and converted to body mass by Barnes et al. (2008) using published regression equations. Only vertebrate prey were identified and measured, with the vast majority being identified to species. -->

<!-- ### Afon Hirnant -->

<!-- The Afon Hirnant is a tributary of the Welsh Dee in North Wales, U.K. (52$^\circ$ 52'N 03$^\circ$ 34) with the study being conducted at three sites. The pH within each site ranged from about 5.5 to 7 over the course of the year. A total of 180 invertebrate samples were collected for the whole sampling periods. All individuals within the benthic samples were counted and identified to the lowest possible taxonomic level (usually species). Prey items were identified using a combination of taxonomic keys and reference slides of previously identified species, after @schmid-arayaConnectanceStreamFood2002; @schmid-arayaImportanceMeiofaunaFood2002. -->

<!-- ### Tadnoll Brook -->

<!-- Description of Tadnoll Brook here -->

<!-- ### Coilaco -->

<!-- Description of Coilaco here -->

<!-- ### Guampoe -->

<!-- Description of Guampoe here -->

<!-- ### Trancura -->

<!-- Description of Trancura here -->

## Allometric Diet Breadth Model (ADBM)

The allometric diet breadth model (ADBM) is based on optimal foraging theory, specifically the contingency foraging model [@macarthurOptimalUsePatchy1966]. We chose this model for various reasons, including because it uses the body size of organisms as input to predict feeding interactions, and because of our familiarity with it. The ADBM predicts the set of prey types (e.g. species or sizes classes) a consumer should feed upon to maximise its rate of energy intake [@petcheySizeForagingFood2008]. The foraging variables in the model are: energy content of the resources, handling times of the prey, space clearance rate, and prey densities. These are allometrically scaled to the body sizes of the species. Further details on the foraging rules defined in the ADBM and ADBM's predictive power across different food webs can be found in @petcheySizeForagingFood2008.

## Assessment of prediction

The accuracy of the predicted diet of the predators was measured using true skill statistic (TSS) which takes into account the true and false predictions of both the presence and absence of links defined as:

$$ \text{TSS} = \frac{ad-bc}{(a+c)(b+d)} $$ where $a$ is the number of observed links that are predicted by the model (true positives), $d$ is the number of observed absences of links that are correctly predicted (true negatives), $b$ is the number of false positives, and $c$ is the number of false negatives. The $TSS$ ranges from $-1$ to $1$, where +1 indicates a perfect prediction. A $TSS$ value of zero or less indicates a performance no better than random [@alloucheAssessingAccuracySpecies2006].

## Inferring food web using partial gut content data

From an empirical dataset of gut content data, we take a random sample of gut contents of specific size (see below) to create a partial gut content dataset. We then fit the ADBM to this partial dataset.

To fit the ADBM to a gut content dataset (be it complete or partial), we used the rejection ABC method we previously developed in @gupta2021 to accept a parameter value from a prior distribution which would have resulted in the minimum distance, where distance = 1 - True skill statistic. The true skill statistic was computed between the diets predicted from the ADBM, and those observed in the sampled gut content data. We repeated this process $n~(= 100)$ times for every $i$ number of guts, where $i$ lies between 1 and total number of gut content data in the pool.

*Input:*

-   Predators $P: P = \{p_1, p_2, \dots, p_k \}$

-   A pool of gut content data $G: G = \{g_1, g_2, \dots, g_n\}$, where $g_{n}$ is the observed diet matrix containing ones and zeros. [OP: I made some edits here, which I'm not sure if are correct.]

-   A model prediction $model(\theta): ADBM(\theta) = \{d_{p_1}, d_{p_1}, \dots, d_{p_k}\}$, where $d_{p_k}$ is the predicted diet matrix of predator $k$ containing ones and zeros.

-   A summary statistic $s(x): s(x) \subseteq model(\theta)$ [OP: a short explanation would be good.]

-   A distance function $d(x_i, y) : d(x_i,y) = 1 - TSS(x_i, y)$  [OP: a short explanation would be good.]

-   An observed food web $Y: Y = \{d_{p_1}', d_{p_1}', \dots, d_{p_k}'\}$  [OP: a short explanation would be good.]

*Sampling:*

for $i = 1, \dots, tgut$ where $tgut$ is the total number of gut content data in the pool $G$

-   for $j = 1, \dots, nsample$ where $nsample$ is the number of independent samples to be drawn

    -   Draw a set of gut content data $y = \{g_1, g_2, \dots, g_i\}$ from the pool of gut content data $G$

    -   for $k = 1, \dots, npar$ where $npar$ is the number of parameter values to be sampled

        -   Draw a set of parameter values $\theta_k$ from the prior distribution $\pi(\theta)$

        -   Compute the model result $x_i = model(\theta_k)$

        -   Compute $s(x_i)$ and $d(s(x_i), y)$

    -   Accept $\theta_j$, which results in the $min_i\{d(s(x_i), y)\}$

-   Compute $TSS_{i}(x, y) = \{TSS(x_i,y): x_i = ADBM(\theta_j), \theta_j \text{ computed from previous step}\}$ using the accepted $\theta_1, \dots, \theta_{nsample}$

*Output:*

The $TSS$ between observed and predicted food webs, and the posterior parameter distributions for every $i$ number of gut content data drawn from the pool of gut content data.

# Results

We first present how the accuracy of the food web model in predicting trophic interactions varies with increasing amount of gut content data provided to the food web model. Then, we calculate the number of gut content data for each food webs that results in accuracy equal to 95% of the maximum true skill statistics predicted by the food web model. We also compute how well  the food web model predicts structural properties from incomplete gut content data. by computing mean standardised error in food web properties.

## Inferring trophic interactions using ADBM and incomplete gut content data

```{r, echo=FALSE, fig.align='center', out.width="500px", fig.cap="\\label{fig:fig_ra} (a, c, e, g, i, k, m) Accuracy of the predicted food web measured using the true skill statistics, predicted by the ADBM parameterised using gut content data. Line and shaded grey region represents the mean and the prediction interval of 100 independent samples respectively. Red line represents the number of predator guts required to achieve a TSS of 95\\% of the maximum TSS. (b, d, f, h, j, l, n) Ratio of the true skill statistic of the predicted food web by the ADBM parameterised using gut content data to that of the predicted food web constructed using gut content data only. The dashed line shows where the ratio is one for reference."}
knitr::include_graphics("../../results/misc/plot_TSS_ratio.png")
```

The true skill statistics of the food webs predicted by the ADBM using subsets of gut content data improved quickly for lower number of predator guts (Fig. \ref{fig:fig_ra}) (a, c, e, g, i, k, m). Furthermore, the prediction interval of the true skill statistics decreased with increasing number of predator guts with the mean TSS asymptoting to the maximum mean TSS achieved by the ADBM when all the gut content data was used. Although the maximum TSS varied among the food webs, the qualitative increase in the TSS was the same.

For Broadstone Stream food web, with only 381 gut contents, which is 38% of the total gut content data, the ADBM predicted the food web with the TSS of 0.74 which was 95% of the mean TSS (0.78) achieved using complete gut content data (Fig. \ref{fig:fig_ra}(a)). In case of the Celtic Sea food web, only 171 gut content data which is 35% of the total gut content data was required by the ADBM to predict food web with TSS equal to 95% of the mean TSS (0.68) achieved using complete gut content data (Fig. \ref{fig:fig_ra}(c)).

For a low number of predator guts, the TSS of the model predicted food web was higher than the TSS of the food web constructed from gut content data only (Fig. \ref{fig:fig_ra} (b, d, f, h, j, l, n)). An increase in the number of predator guts did reduce the ratio TSS(Model)/TSS(data) reduced to less than one, and reduced further gradually. This shows that with less data available it is better to use the model, while if more data is available then it is better to use the data itself.

## Mean standardised error in structural properties with respect to the number of gut content data

Mean standardised error in structural properties of Afon Hirnant, Broadstone Stream, Celtic Sea and Tadnoll Brook food web reduced drastically with increasing number of predator gut content data at low number of predator gut content data (Fig. \ref{fig:fig_rab}). In case of Broadstone Stream food web, the 95% prediction interval reduced as well with increasing number of predator gut content data.

\*\*AG: Working on this paragraph.

```{r, echo=FALSE, fig.align='center', out.width="500px", fig.cap="\\label{fig:fig_rab} Mean standardised error in structural properties of different food webs with respect to the number of gut content data."}
knitr::include_graphics("../../results/misc/plot_mse_ngut_all.png")
```

## Dependence of the minimum gut content data on number of species, links and connectance

There was a positive trend between minimum gut content data [OP: minimum that could be used in order to ensure 95% of maximum TSS???] and number of species (S) (Fig. \ref{fig:fig_rb} (a)), links (L) (Fig. \ref{fig:fig_rb} (c)), connectance (C) (Fig. \ref{fig:fig_rb} (d)) and number of predator nodes (Fig. \ref{fig:fig_rb} (e)), with Broadstone Stream being an outlier in Fig. \ref{fig:fig_rb} (a, c, e). The minimum gut content data increases with number of gut content data sampled (Fig. \ref{fig:fig_rb} (b)). Furthermore, there seems to be a negative relationship between proportion of minimum gut (= Number of minimum gut / Number of maximum guts sampled) and number of species (S) (Fig. \ref{fig:fig_rb} (f)), links (L) (Fig. \ref{fig:fig_rb} (g)), connectance (C) (Fig. \ref{fig:fig_rb} (h)).

```{r echo=FALSE, fig.align='center', out.width="500px", fig.cap="\\label{fig:fig_rb} (a, b, c, d, e) Number of minimum gut content data plotted against number of species (S), number of maximum guts sampled, number of links (L), connectance (C) and number of predator nodes respectively. (f, g, h) Proportion of minimum gut content data plotted against number of species (S), number of links (L) and connectance (C) respectively."}
knitr::include_graphics("../../results/misc/plot_min_gut_vs_X.png")
```

```{r echo=FALSE, fig.align='center', out.width="500px", fig.cap="\\label{fig:fig_rc} Link accumulation curve of the food web constructed using gut content data plotted against number of gut content data."}
knitr::include_graphics("../../results/misc/plot_emplinks_ngut.png")
```

# Discussion

We have demonstrated how a food web model can be used to predict food webs reasonably well even when incomplete data about trophic interactions is available. At lower amount of gut content data, predicting with the model is better than using the data itself, while at higher amounts of gut content data, it is better to only use the data. We computed the upper limit of amount of gut content data till when it can be used to compensate for undersampling [OP: reword this sentence, it is unclear.]. This can help inform how much gut content data to actually collect when we are using a food web model to infer trophic interactions. A future development could be to make the same assessment using other food web models, and to also use food web data other than gut content data to parameterise those models.

## Why incomplete gut content data can be used to infer trophic interactions?

In all of the seven food webs, the ADBM was able to infer the trophic interactions using incomplete gut content data because the presence absence information from the gut content data was still sufficient to constrain the possible model parameter values of the ADBM that best explained the predators' diets. Although in theory the ADBM can predict trophic interactions using only body sizes of organisms as it is based on set of foraging rule, it still requires some presence absence data to constrain the posterior parameter space thereby making more accurate predictions.

## Why does the minimum gut data differ among food webs?

We found the minimum amount of gut content data to vary across food webs and this could be due to the possibility that the food webs are sampled differently, i.e. some food webs are more extensively sampled than others (Fig. \ref{fig:fig_rc}). Furthermore, it is possible that interactions in some of the food webs are are more size-structured than others, and are therefore better predicted by the ADBM and consequently require a lower number of gut content data for model parameterisation.

\*\*:AG: I would need to add information about the interactions in these food webs.

## Model fills up missing interactions and guides sampling direction

When little data about trophic interactions is available, the model can be used to add data about interactions for which data is missing. Furthermore, the model predictions can be used as a guide to sample predator guts that might be significant in the food web [OP: how? I suggest to explain this in more detail.]

## Why is model's TSS less than one?

The food web model used (the ADBM) cannot explain all the interactions in any observed food web that it has been fit to. The foraging rules it encodes are based on the body size and have particular structure and assumptions; not all of these are met by all observed interactions. For example, the ADBM can only predict diets that are contiguous with respect to the size of prey. I.e. it cannot predict that a predator will consume taxa of size 1 and 3, and not taxa of size 2. Hence, if the observed diets are not contiguous when prey are ordered by their size, the estimation process could lead to a lower value of the TSS.

Furthermore, the observed data may be missing links, e.g. links that rarely occur. Hence the ADBM's maximum predicted true skill statistics is less than one for all food webs it has so far been fit to.

## TSS(Model)/TSS(Data) for different food webs

\*\*AG: Working on this.

## Why Broadstone Stream is an outlier?

Broadstone Stream is the most sampled food web compared to all the other food webs (Fig. \ref{fig:fig_rc}), which means the observed Broadstone Stream food web might be more accurate compared to others.

\*\*AG: Working on this, not sure if this paragraph is needed.

## Propagating uncertainty from gut content data to model's predictions

\*\*OP Focus on this, at least in the context of further work that could be done. ## There is uncertainty in gut content data and needs to be propagated into model's predictions

There is considerable uncertainty involved in gut content analysis [@bakerFishGutContent2014] such as in fish's guts, there are sometimes loose tissues that are not identifiable and cannot be assigned to a specific prey item with certainty. There are factors such as sample size of consumers, mechanical prey handling, differential digestion and evacuation rates of different prey types and volumes, and the ingestion order that in combination result in an unquantifiable error which is difficult to interpret in the predator diet [@hyslopStomachContentsAnalysis1980; @rindorfBiasEstimatingFood2004; @bakerFishGutContent2014]. Therefore, a future development would be to consider how this uncertainty is propagated into food web prediction.

## Comparing our study with other studies of predicting food webs using incomplete data

[OP: This one I already shared with you: https://onlinelibrary.wiley.com/doi/10.1111/ele.13966

\*\*OP: Focus on this, and other studies that have looked at the importance of quantity and quality of food web data ## Comparing our study with other studies of predicting food webs using incomplete data

Some studies have presented how the accuracy of food web prediction change when we vary the amount of food web data [@grayJoiningDotsAutomated2015]. 
\*\*AG: Working on this paragraph.

## What is the generalisability of our results?

\*\*OP: Focus on this.

In our study, we have implemented the approach only with the ADBM. However, in principle, this approach could be extended to other food web models [OP: such as???]. A future prospect could be to study how well different food web models' prediction accuracy varies for different amount of gut content data. This can also help in making decision as to which food web model to chose from given a set of gut content data.

[OP: seems to not add much relative to the previous paragraph.] A future study could be to use our approach of estimating the minimum gut content data for different food web models. An efficient model would be the one that uses minimum available gut content data to predict species interactions with high accuracy and high precision.

[OP: comment if and why one might expect other food web models to show different efficiency.]

## Assumption that the food web constructed with the total gut content data is the true food web and how can this be improved

[OP: this paragraph / section needs further development]

Some of the food webs are not very well predicted using the ADBM. And with the assumption that the food web constructed from the complete gut content data is the true food web might result in not so accurate estimates of the minimum gut content data. As it might be the case that the links are still under-sampled for many nodes [@woodwardIndividualBasedFoodWebs2010].

## How can we include other types of food web data?

\*\*OP: Focus on this ## How can we include other food web data type?

A future prospect would be to include other types of food web data along with gut content simultaneously to parameterise the ADBM. For example, one could use the approximate trophic position inferred from stable isotope ratio data from a tissue of an individual and gut content data of a different predator simultaneously to parameterise a food web model such as the ADBM. In this case, trophic position would be an additional summary statistic required in the parameterisation method. This combination of different types of data about trophic interactions could be particularly useful when only a limited amount of diet information can be extracted by using a single method.

# Conclusion

\*\*AG: Working on this.

# Supplementary Information

<!-- ```{r, echo=FALSE, fig.align='center', out.width="400px", fig.cap="\\label{fig:fig_s1} \\textit{In silico} approach for investigating the effect of quantity and composition of food web data in predicting food web structure using the ADBM."} -->

<!-- knitr::include_graphics("../fig/flowchart.png") -->

<!-- ``` -->

## Gut content data simulation

\*\*OP: It needs to be very clear in the main text, and here, why this is being described. It is currently unclear, not least because the main text does not, I think, refer to the information here. [OP: I'm not sure my comment has been addressed. We should discuss this.] [OP: I don't see this Supplement being referred to anywhere in the main text. I would expect it to be.] [OP: I advise to integrate the content of this supplement into the main article. There is not so much of it, and it is insightful.]

We simulated a food web using the ADBM for a given set of parameters. For a given set of predators, we subset the diet from the simulated food web. Then using a probability mass function (distribution), we sampled the gut content data from predators' diet thereby incorporating the uncertainty in the gut content data. We repeated this process multiple number of times for every predator in the food web.

*Input:*

-   Predators whose diet are to be simulated $P = \{p_1, \dots, p_n\}$

-   A simulated food web $ADBM(\theta_i) = \{d_{p_1}, d_{p_1}, \dots, d_{p_k}\}$, where $d_{p_k}$ is a one-dimensional diet matrix of predator $k$ containing ones and zeros.

-   A function which describes uncertainty in the diet $U(d)$

-   Number of independent guts to be simulated for a predator $p_i: ngut$

*Sampling:*

-   for $p_i \in P$

    -   for $j = 1, \dots, ngut$, where $ngut$ is the number of guts to be simulated

        -   Simulate a single gut of a predator $p_i: g(p_i) = d_{p_i} * U(d_{p_i})$

    -   Set of gut of a predator $p_i: G(p_i) = \{ g(p_i) : g(p_i) = d_{p_i}*U(d_{p_i}) \}$

*Output:*

-   We simulated a pool of gut content data which contains simulated gut content data $G(p_i)$ for every predator $p_i$

## Prediction using gut content data from a simulated food web

```{r fig_r0, echo=FALSE, fig.align='center', out.width="400px", fig.cap="\\label{fig:fig_r0} True skill statistics between predicted food web and observed food web for a simulated small reef food web estimated for distinct predator guts in a sample. The observed simulated food web consists of 50 species and \\dots links. The vertical bars correspond to the prediction intervals of the true skill statistics with filled circles representing the corresponding mean. A prediction interval of the TSS is formed using a set of 100 accepted TSS values using the ABC method."}
knitr::include_graphics("../fig/TSS_with_n_pred_prop_simulated.png")
```

The true skill statistics (TSS) between the predicted food web and simulated observed food web saturated with an increasing number of distinct predator guts in a sample (Fig. \ref{fig:fig_r0}). The TSS of the predicted food webs estimated using the complete individual gut data had narrower prediction intervals resulting in less uncertainty, and higher mean TSS than that using the partial individual gut data. The maximum limit of the prediction interval of TSS estimated using the complete gut data and the partial gut data were almost equal, with the minimum limit of the prediction interval of TSS using partial gut data being lower than that from the complete gut data. Eventually, the gap between the mean TSS using the partial gut data and the complete gut data reduced with an increasing number of distinct predator guts. This suggests that when there is enough predator species' gut data, the achieved TSS was almost constant and hence independent of the amount of gut data.

The maximum TSS estimated using the complete gut data was very close to one and almost remained constant with an increasing number of different predator species sampled. With the gut data sample of only five distinct predator species, 95% of the maximum mean TSS was achieved when complete individual gut data was used, while the same was achieved with 15 predator species for partial gut data. This shows that one does not need to know the gut data of all the species to predict the food web and the accuracy is dependent on the completeness of an individual gut data.

```{r fig_r3, echo=FALSE, fig.align='center', out.width="300px", fig.cap="\\label{fig:fig_r3} True skill statistics between predicted food web and observed food web estimated for different number of distinct predator guts. The estimation is done for three sets of gut data: gut content data of predators whose body sizes are smaller than the mean body size, larger than the mean body size, and all the gut content data."}
knitr::include_graphics("../fig/TSS_with_n_pred_prop.png")
```

```{r fig_r31, echo=FALSE, fig.align='center', out.width="500px", fig.cap="\\label{fig:fig_r31} True skill statistic between predicted food web using ADBM and incomplete gut content data, and observed food web against the true skill statistic between food web constructed using incomplete gut content data, and observed food web. Error bars represent prediction intervals of 100 independent samples. Dashed line is 1:1 line for reference."}
knitr::include_graphics("../../results/misc/TSS_model_vs_data.png")
```

```{r fig_r32, echo=FALSE, fig.align='center', out.width="500px", fig.cap="\\label{fig:fig_r32} Mean standardised error in structural properties in the food web predicted using ADBM and incomplete gut content data against structural properties in the predicted food web constructed using incomplete gut content data. Error bars represent prediction intervals of 100 independent samples. Dashed line is 1:1 line for reference."}
knitr::include_graphics("../../results/misc/mse_model_vs_data.png")
```

<!-- # Notes: -->

<!-- ## Stable isotope ratios -->

<!-- * Isotope ratios are reported in parts per thousand  relative to standards, and atmospheric N for nitrogen, and defined in delta notation as -->

<!-- * $\delta^{15}N$ or $\delta^{13}C = (\frac{R_{sample}}{R_{standard}} - 1) * 10^3 $ -->

<!-- * Most commonly a fixed $\Delta^{15}N$ value of 3.4 ppt is used to estimate relative species TP -->

<!-- # Other notes: -->

<!-- * Audience: Researchers who are interested in food web data collection -->

# Acknowledgements

# Author contributions

# References

<!-- ## Why knowing trophic interactions is important? -->

<!-- Knowledge about the trophic interactions in a food web can be used to answer crucial ecological questions: Which species is a keystone species in a food web [@jordanKeystoneSpeciesFood2009]? How would a food web respond if the top predators are removed [@knightTrophicCascadesEcosystems2005]?  Which species should be given priority when conservation efforts are implemented [@mcdonald-maddenUsingFoodwebTheory2016]? How robust is a food web to extinctions [@dunneNetworkStructureBiodiversity2002]? -->

<!-- ## What are the methods for knowing interactions? -->

<!-- Species interactions can be directly or indirectly inferred from different types of methods such as gut content analysis [@peralta-maraverStructureDynamicsStability2017], stable isotope ratio analysis of tissues [@laymanCanStableIsotope2007], experimentation [@warrenSpatialTemporalVariation1989], DNA metabarcoding [@roslinUseDNABarcodes2016], food web model [@gravelInferringFoodWeb2013; @petcheySizeForagingFood2008; @williamsSimpleRulesYield2000] and literature research [@grayJoiningDotsAutomated2015; @cohenSoilInvertebratesChemistry2014; @goldwasserConstructionAnalysisLarge1993a]. -->

<!-- ## What are the shortcomings of the methods mentioned above? -->

<!-- These food web methods mentioned above have multiple shortcomings. Stable isotope ratio analysis of the organism's tissue does not give a direct information of the diet of that organism but rather approximate trophic position of that species in the food web. Although mixing models can be used to determine what prey items are most likely fed upon by a predator but this results in uncertainty in the estimates [@kadoyaIsoWebBayesianIsotope2012; @crawfordApplicationsStableIsotope2008]. Experimentation (e.g. feeding trials) may create unrealistic conditions where a predator is confined in a small space with a potential "prey" and is effectively "forced" into eating it (citreqd). DNA metabarcoding could have many issues such as failure to relate OTUs to species, or inability to distinguish secondary predation, i.e. the DNA signature of the prey consumed by the prey consumed by the primary item found in the predator's gut. Another approach to infer missing links is literature research which involves assuming that a link will occur because another author described it in some other system, or for related predator/prey species, e.g. from the same genus or family  [@grayJoiningDotsAutomated2015; @cohenSoilInvertebratesChemistry2014; @goldwasserConstructionAnalysisLarge1993a] (citreqd). This typically results in a food web that will be very oversampled with false positives in the presence of links (citreqd). -->

<!-- ## Food web model -->

<!-- Food web models are set of rules governed by ecological theory which defines trophic interactions in a food web. Along with inferring missing links in a food web, they can be used for understanding the mechanistic of the food web and ecological forecasting. Observed food web data can be used to parameterise the food web model. -->

<!-- ## Parameterising a food web model using food web data -->

<!-- One of the common method of knowing presence-absence information of trophic links in food web is by gut content analysis of predator guts, which can be used to parameterise a food web model. Gut content data have been used to predict the diet of a predator in a food web [@eitzingerTestingValidityFunctional2018; @omalleyEffectsFoodWeb2017; @dixonAssessingDietNorth2017; @mcclain-countsTrophicStructureMesopelagic2017b; @peralta-maraverStructureDynamicsStability2017]. A single gut content data of a predator gives the information what the predator has consumed recently. The diet of that individual can be inferred from the items in the stomach. -->

<!-- ## Why inferring interactions is not easy? -->

<!-- As a food web can be quite complex with many species and trophic interactions [@williamsSimpleRulesYield2000;@berlowSimplePredictionInteraction2009], inferring these trophic interactions with high accuracy and low uncertainty would require a large amount of food web data (citreqd). E.g. One would need to sample many individuals for gut content data, and processing these guts require a large amount of effort and time as well. -->

<!-- A single gut content of a predator provides information of the prey individuals that have been consumed by the predator in a small temporal and spatial scale (citreqd). Therfore, given a spatial and temporal scale of a food web, the sampling of the gut content data needs to be carried out. This means that for a food web constructed on a large spatial and temporal scale large number of gut content data needs to be collected. There is also a possibility that some of the information provided by these types of food web data are redundant i.e. taking into account any further food web data does not necessarily increase the food web prediction significantly. E.g. This can happen if the guts of two predator individuals give the same information about the prey the predator has consumed. This issue can be resolved by using a yield-effort curve which suggests when sampling of new predator guts can be stopped [@hernandezParasitesAlterTopology2008]. -->

<!-- ## Undersampled food web -->

<!-- When it is difficult to obtain food web data, this can result in missing links in the food web thereby resulting in an undersampled food web. Limitations of resources and time can also lead to an undersampled food web because collecting and analysing gut content data is tedious and expensive task (citreqd). It is crucial to have well sampled food web because food web properties are highly sensitive to varying levels of sampling effort [@goldwasserSamplingEffectsEstimation1997]. So, it is important to predict food web properties with high accuracy and high precision. -->

<!-- ## What is a food web model and how can it be used to compensate for undersampling? -->

<!-- When we have an undersampled food web, food web models can be used to predict which trophic interactions are feasible and which are not [@allesinaGeneralModelFood2008; @cohenStochasticTheoryCommunity1985; @gravelInferringFoodWeb2013; @petcheySizeForagingFood2008; @tamaddoni-nezhadConstructionValidationFood2013]. Food web model are set of rules governed by ecological theory which defines trophic interactions in a food web. For example, @petcheySizeForagingFood2008 based on foraging traits allometrically scaled using body size data of species predicted trophic interactions where the model was parameterised using known presence and absence of trophic interactions. -->

<!-- ## Research question -->

<!-- How much gut content data is required when a food web model is used to predict trophic interactions? We are interested in maximizing the accuracy and precision of the predicted food web and simultaneously minimizing the amount of gut content data used, because we do not want to spend a lot of resources in sampling gut content data. -->