### 2022-06-20 inst/report/sampling_standardization/sampling_standardisation_ngut.Rmd
Incorporated most comments from Guy, Eoin and David

---
title: "sampling_standardisation_ngut"
author: "Anubhav Gupta"
date: '2022-05-09'
output: 
  html_document:
    fig_caption: true
    number_sections: true
---

We standardise sampling across food webs.

## Link accumulation curve with respect to the number of predator guts

```{r echo = FALSE, fig.align='center', out.width="600px", fig.cap="Number of trophic links plotted against the number of gut content data for seven food webs."}
knitr::include_graphics("../../../results/misc/plot_emplinks_ngut.png")
```

## Rolling average of link accumulation curve with respect to the number of predator guts

```{r echo = FALSE, fig.align='center', out.width="600px", fig.cap="Rolling average (with window of 30 items) of number of trophic links plotted against the number of predator guts for seven food webs."}
knitr::include_graphics("../../../results/misc/plot_emplinks_ngut_ra.png")
```

## Slope of link accumulation curve with respect to the number of predator guts

```{r echo = FALSE, fig.align='center', out.width="600px", fig.cap="Slope of link accumulation curve plotted against the number of predator guts for seven food webs."}
knitr::include_graphics("../../../results/misc/plot_slope_emplinks_ngut.png")
```

## Rolling average of slope of link accumulation curve with respect to the number of predator guts

```{r echo = FALSE, fig.align='center', out.width="600px", fig.cap="Rolling average (with window of 10 items) of slope of link accumulation curve plotted against the number of predator guts for seven food webs."}
knitr::include_graphics("../../../results/misc/plot_slope_emplinks_ngut_ra.png")
```

## Rolling average of slope of link accumulation curve with respect to the number of predator guts for the Broadstone Stream food web

```{r echo = FALSE, fig.align='center', out.width="600px", fig.cap="Rolling average (with window of 100 items) of slope of link accumulation curve plotted against the number of predator guts for seven food webs."}
knitr::include_graphics("../../../results/misc/plot_slope_emplinks_ngut_ra_bs.png")
```

## Factor

```{r echo = FALSE}
factor_all <- readRDS("../../../results/misc/factor_all_ngut.RDS")
print(factor_all)
```

## Corrected minimum number of gut vs number of species

```{r echo = FALSE, fig.align='center', out.width="600px", fig.cap="Corrected minimum number of gut vs number of species (S) for seven food webs."}
knitr::include_graphics("../../../results/misc/plot_corrected_mingut_vs_S_ngut.png")
```

## Corrected minimum number of gut vs number of links

```{r echo = FALSE, fig.align='center', out.width="600px", fig.cap="Corrected minimum number of gut vs number of links (L) for seven food webs."}
knitr::include_graphics("../../../results/misc/plot_corrected_mingut_vs_L_ngut.png")
```

## Corrected minimum number of gut vs connectance

```{r echo = FALSE, fig.align='center', out.width="600px", fig.cap="Corrected minimum number of gut vs connectance (C) for seven food webs."}
knitr::include_graphics("../../../results/misc/plot_corrected_mingut_vs_C_ngut.png")
```